No Evidence for Lunar Transit in New Analysis of HST Observations of the Kepler-1625 System

### 2019-02-13 arxiv/ms.tex
ready to submit

\submitjournal{AJ}

@@

\begin{abstract}
    Observations of the Kepler-1625 system with the Kepler and Hubble Space Telescopes have suggested the presence of a candidate exomoon, Kepler-1625b I, a Neptune-radius satellite orbiting a long-period Jovian planet. Here we present a new analysis of the Hubble observations, using an independent data reduction pipeline. We find that the transit light curve is well fit with a planet-only model, with a best-fit $\chi^2_\nu$ equal to $1.01$. The addition of a moon does not significantly improve the fit quality. We compare our results directly with the original light curve from \cite{teachey18b}, and find that we obtain a better fit to the data using a model with fewer free parameters (no moon).  We discuss possible sources for the discrepancy in our results, and conclude that the lunar transit signal found by \cite{teachey18b} was likely an artifact of the data reduction. 
\end{abstract}

@@

\begin{figure}
\includegraphics[width = 0.5 \textwidth]{figures/fig4_shifts.pdf}
    \caption{Shift (in pixels) relative to the mean position of the spectrum in the spectral direction (top) and spatial direction (bottom). The largest shift occurs after orbit 14 due to a guide star reacquisition.}
\label{fig:shifts}
\end{figure}

@@

\begin{figure}
\includegraphics[width = 0.5 \textwidth]{figures/fig6_rawdata.pdf}
    \caption{Top: extractred light curve from this work (blue) compared to TK18 (red). Bottom: the difference in photoelectron counts between the data sets. The moon ingress identified by TK18 is marked by the dotted gray line.}
\label{fig:raw}
\end{figure}

@@

\begin{figure*}
\includegraphics[width = 1.0 \textwidth]{figures/fig3_bestfits.pdf}
    \caption{Best fit models compared to normalized transit light curves from this work (left, blue) and from TK18 (right, red). The top panel shows the best fit no-moon model, and the bottom shows the best fit moon model. The lower left of each panel indicates the fit rms (in ppm), the $\chi_2$, the degrees of freedom, and the BIC.  Each light curve is divided by its best fit systematics model (XY decorrelation for this work; exponential and offset for TK18).  The dotted gray line marks the possible moon ingress identified by TK18.}
\label{fig:bestfit}
\end{figure*}

@@

\begin{figure}
\includegraphics[width = 0.5 \textwidth]{figures/fig2_rms.pdf}
    \caption{Light curve rms versus bin size for the best fit no-moon model (solid lines) and moon model (dashed lines), for data from this work (red) and TK18 (blue). The fits to data from this work agree well with the expected photon-limited, $\sqrt{N}$ decrease in rms with bin size (black line).  We also reach the photon limit for the TK18 data, but only for the moon model. The rms for the no-moon fit (red dashed line) ranges from $1.1 - 1.5\times$ the photon limit for 1 to 20 points per bin.}
 \label{fig:rms}
\end{figure}

@@

\subsection{Comparison with Teachey \& Kipping (2018)}
TK18 found evidence for the transit of a Neptune-size moon in their analysis of the HST data, in contrast to the findings presented here.  To make a direct comparison with their results, we fit the TK18 light curve directly.  We fit the astrophysical signal with both the no-moon and models, and used the exponential systematics model.  Figure\,\ref{fig:bestfit} shows the best fit models.

@@

Similar to the findings of TK18, the moon model improved the fit quality by a $\Delta\chi^2 = 29.9$. Notably, however, the moon model fit to the TK18 data does not perform better than the no-moon model fit to our new data (rms of 362 versus 356 ppm), even with the additional seven free parameters.  %This difference in fit quality corresponds to $\Delta\mathrm{BIC} = 31.2$, which is strong evidence against the moon.

@@

\begin{figure}
\includegraphics[width = 0.5 \textwidth]{figures/fig5_pairs.pdf}
    \caption{Posterior distributions for the moon radius and time of transit based on an MCMC fit to data from this work (blue) and from TK18 (red). The shading corresponds to $1-$, $2-$, and $3\,\sigma$ contours (from darkest to lightest). These values are marginalized over all other model parameters.} 
\label{fig:corner}
\end{figure}

\section{Discussion}
A natural question arising from our results is what the difference is between TK18 and the new fits presented here. Since we found that

One possibility is that the source of the discrepancy is the extracted transit light curves themselves. As shown in Figure\,\ref{fig:raw}, our measured count rate ranges from $2.46 -2.74\%$ lower than the TK18 light curve, and there is a small bump in the difference between the two data sets near the location of the moon transit identified in the TK18 data (see the bottom panel).

 We explored several modifications to our pipeline to attempt to reproduce the TK18 reduction. These included rotating the image by 0.5 degrees, using the same aperture as TK18 to extract the spectrum, and scaling the master sky flat for the background subtraction (rather than just subtracting the median). None of these modifications had a significant effect on our results.

@@

@@

In addition to these differences in data reduction, we also used an alternative detrending model. We opted for a decorrelation against X and Y position, motivated by known variation in the sensitivity of individual pixels on the detector and the presence of small shifts in the position of the spectrum (see Figure\,\ref{fig:shifts}). We applied this detrending model to the TK18 data, and obtained a poor fit to the light curve compared to exponential detrending (for the moon model with XY decorrelation, the rms was 404 ppm). We did not recover a large moon in this fit (the best fit was $r_\mathrm{moon} = 1.2R_\oplus$), but there was correlated noise present in the data that precluded a robust measurement of the uncertainty on the moon radius for this model. The poor quality of the fit points toward the extracted light curve itself as the culprit for the difference in our results, but as we discuss, we could not identify a specific step in the data reduction that is responsible.

 We conclude that no single choice in the data reduction or the modeling provides an easy explanation for the difference in our results, but rather that the difference is more likely due to the combination of many subtle choices.

\section{Conclusion}
The hunt for exomoons remains exciting and volatile. Our analysis of the exomoon candidate Kepler-1625b I shows that the evidence in favor of the moon is highly sensitive to details of the data reduction and modeling. Given that we achieve an excellent fit to the light curve with a no-moon model, and that the addition of a moon does not significantly improve the fit quality, we suggest that future allocation of follow-up resources for exomoons should prioritize alternate targets.

>>>>

\submitjournal{ApJL}

@@

\begin{abstract}
    Observations of the Kepler-1625 system with the Kepler and Hubble Space Telescopes have suggested the presence of a candidate exomoon, Kepler-1625b I, a Neptune-radius satellite orbiting a long-period Jovian planet. Here we present a new analysis of the Hubble observations, using an independent data reduction pipeline. We find that the transit light curve is well fit with a planet-only model, with a best-fit $\chi^2_\nu$ equal to $1.01$. The addition of a moon does not significantly improve the fit quality. We compare our results directly with the original light curve from \cite{teachey18b}, and find that we obtain a better fit to the data using a model with fewer free parameters (no moon).  We discuss possible sources for the discrepancy in our results, and conclude that the lunar transit signal found by \cite{teachey18b} was likely an artifact of the data reduction.  This finding highlights the need to develop independent pipelines to confirm results that push the limits of measurement precision.
\end{abstract}

@@

\begin{figure}
\includegraphics[width = 0.5 \textwidth]{fig1_shifts.pdf}
    \caption{Shift (in pixels) relative to the mean position of the spectrum in the spectral direction (top) and spatial direction (bottom). The largest shift occurs after orbit 14 due to a guide star reacquisition.}
\label{fig:shifts}
\end{figure}

@@

\begin{figure}
\includegraphics[width = 0.5 \textwidth]{fig2_rawdata.pdf}
    \caption{Top: extractred light curve from this work (blue) compared to TK18 (red). Bottom: the difference in photoelectron counts between the data sets. The moon ingress identified by TK18 is marked by the dotted gray line.}
\label{fig:raw}
\end{figure}

@@

\begin{figure*}
\includegraphics[width = 1.0 \textwidth]{fig3_bestfits.pdf}
    \caption{Best fit models compared to normalized transit light curves from this work (left, blue) and from TK18 (right, red). The top panel shows the best fit no-moon model, and the bottom shows the best fit moon model. The lower left of each panel indicates the fit rms (in ppm), the $\chi_2$, the degrees of freedom, and the BIC.  Each light curve is divided by its best fit systematics model (XY decorrelation for this work; exponential and offset for TK18).  The dotted gray line marks the possible moon ingress identified by TK18.}
\label{fig:bestfit}
\end{figure*}

@@

\begin{figure}
\includegraphics[width = 0.5 \textwidth]{fig4_rms.pdf}
    \caption{Light curve rms versus bin size for the best fit no-moon model (solid lines) and moon model (dashed lines), for data from this work (red) and TK18 (blue). The fits to data from this work agree well with the expected photon-limited, $\sqrt{N}$ decrease in rms with bin size (black line).  We also reach the photon limit for the TK18 data, but only for the moon model. The rms for the no-moon fit (red dashed line) ranges from $1.1 - 1.5\times$ the photon limit for 1 to 20 points per bin.}
 \label{fig:rms}
\end{figure}

@@

\subsection{Comparison with Teachey \& Kipping (2018)}
TK18 found evidence for the transit of a Neptune-size moon in their analysis of the HST data, in contrast to the findings presented here.  To compare our results with theirs, we fit the TK18 light curve directly.  We fit the astrophysical signal with both the no-moon and models, and used the exponential systematics model.  Figure\,\ref{fig:bestfit} shows the best fit models.

@@

Similar to the findings of TK18, the moon model improved the fit quality by a $\Delta\chi^2 = 29.9$. Notably, however, the moon model fit to the TK18 data does not perform better than the no-moon model fit to our new data (rms of 362 versus 356 ppm), even with the additional seven free parameters.

@@

\begin{figure}
\includegraphics[width = 0.5 \textwidth]{fig5_pairs.pdf}
    \caption{Posterior distributions for the moon radius and time of transit based on an MCMC fit to data from this work (blue) and from TK18 (red). The shading corresponds to $1-$, $2-$, and $3\,\sigma$ contours (from darkest to lightest). These values are marginalized over all other model parameters.} 
\label{fig:corner}
\end{figure}

\section{Discussion \& Conclusions}
A natural question arising from our analysis is the source of the discrepancy between TK18 and the new results presented here. We find that with our new data, there is strong evidence against the moon ($\Delta\rm{BIC} > 25$), even when we use a comparable model to TK18 (a 6-parameter moon model and an exponential fit to the visit-long trend).

If the model is not the source of the difference, it must be the extracted transit light curves themselves.  Figure\,\ref{fig:raw} shows a direct comparison of the light curves. The count rate we measure is $2.46 -2.74\%$ lower than the TK18 light curve, and there is a small bump in the difference between the two data sets near the location of the moon transit identified in the TK18 data (see the bottom panel). This bump may be the source of the moon feature reported in TK18.

 We explored several modifications to our pipeline to attempt to reproduce the TK18 data reduction. These included rotating the image by 0.5 degrees, using the same aperture as TK18 to extract the spectrum, and scaling the master sky flat for the background subtraction (rather than just subtracting the median). None of these modifications had a significant effect on our results.

@@

We conclude that no single choice in the data reduction provides an easy explanation for the difference in our light curves. This finding illustrates the challenge of pushing measurement precision to the 100 ppm level, and highlights the importance of developing multiple independent pipelines to confirm cutting-edge results.

@@

%In addition to these differences in data reduction, we also used an alternative detrending model. We opted for a decorrelation against X and Y position, motivated by known variation in the sensitivity of individual pixels on the detector and the presence of small shifts in the position of the spectrum (see Figure\,\ref{fig:shifts}). We applied this detrending model to the TK18 data, and obtained a poor fit to the light curve compared to exponential detrending (for the moon model with XY decorrelation, the rms was 404 ppm). We did not recover a large moon in this fit (the best fit was $r_\mathrm{moon} = 1.2R_\oplus$), but there was correlated noise present in the data that precluded a robust measurement of the uncertainty on the moon radius for this model. The poor quality of the fit points toward the extracted light curve itself as the culprit for the difference in our results, but as we discuss, we could not identify a specific step in the data reduction that is responsible.

%\section{Conclusion}
%The hunt for exomoons remains exciting and volatile. Our analysis of the exomoon candidate Kepler-1625b I shows that the evidence in favor of the moon is highly sensitive to details of the data reduction. Given that we achieve an excellent fit to the light curve with a no-moon model, and that the addition of a moon does not significantly improve the fit quality, we suggest that future allocation of follow-up resources for exomoons should prioritize alternate targets.

### 2019-02-13 ms-temp.tex
ready to submit

\begin{abstract}
    Observations of the Kepler-1625 system with the Kepler and Hubble Space Telescopes have suggested the presence of a candidate exomoon, Kepler-1625b I, a Neptune-radius satellite orbiting a long-period Jovian planet. Here we present a new analysis of the Hubble observations, using an independent data reduction pipeline. We find that the transit light curve is well fit with a planet-only model, with a best-fit $\chi^2_\nu$ equal to $1.01$. The addition of a moon does not significantly improve the fit quality. We compare our results directly with the original light curve from \cite{teachey18b}, and find that we obtain a better fit to the data using a model with fewer free parameters (no moon). We conclude that the evidence for the transit of Kepler-1625b I hinges on subtleties of the data reduction, and we therefore suggest that future follow-up efforts for exomoons should prioritize alternate targets.

\end{abstract}

@@

\subsection{Comparison with Teachey \& Kipping (2018)}
TK18 found evidence for the transit of a Neptune-size moon in their analysis of the HST data, in contrast to the findings presented here.  To make a direct comparison with their results, we fit the TK18 light curve directly.

We fit the data with both the no-moon and models.  For the systematics, we tested the XY decorrelation model described above, but found that it did not perform as well as the systematics models presented in TK18 (10\% higher rms and time-correlated residuals).  We therefore opted to use a TK18 systematics model to enable a fair comparison between our results. We obtained the best fit using an exponential trend in time, a constant offset after orbit 14 (where the guide stars are reacquired), and the non-parameteric orbit-long ramp model. This model has one more free parameter than the systematics model given in Equation~\ref{eq:sys}. Figure\,\ref{fig:bestfit} shows the best fit models.

Similar to the findings of TK18, the moon model improved the fit quality by a $\Delta\chi^2 = 29.9$. Notably, however, the moon model fit to the TK18 data does not perform better than the no-moon model fit to our new data (rms of 362 versus 356 ppm), even with the additional seven free parameters.  %This difference in fit quality corresponds to $\Delta\mathrm{BIC} = 31.2$, which is strong evidence against the moon.

@@

\section{Discussion}
A natural question arising from our results is what the difference is between TK18 and the new fits presented here. One possibility is that the source of the discrepancy is the extracted transit light curves themselves. As shown in Figure\,\ref{fig:raw}, our measured count rate ranges from $2.46 -2.74\%$ lower than the TK18 light curve, and there is a small bump in the difference between the two data sets near the location of the moon transit identified in the TK18 data (see the bottom panel).

 We explored several modifications to our pipeline to attempt to reproduce the TK18 reduction. These included rotating the image by 0.5 degrees, using the same aperture as TK18 to extract the spectrum, and scaling the master sky flat for the background subtraction (rather than just subtracting the median). None of these modifications had a significant effect on our results.

@@

In addition to these differences in data reduction, we also used an alternative detrending model. We opted for a decorrelation against X and Y position, motivated by known variation in the sensitivity of individual pixels on the detector and the presence of small shifts in the position of the spectrum (see Figure\,\ref{fig:shifts}). We applied this detrending model to the TK18 data, and obtained a poor fit to the light curve compared to the polynomial detrending (for the moon model with X-Y decorrelation, the rms was 404 ppm). We did not recover a large moon in this fit (the best fit was $r_\mathrm{moon} = 1.2R_\oplus$), but there was correlated noise present in the data that precluded a robust measurement of the uncertainty on the moon radius for this model. The poor quality of the fit points toward the extracted light curve itself as the culprit for the difference in our results, but as we discuss, we could not identify a specific step in the data reduction that is responsible.

 We conclude that no single choice in the data reduction or the modeling provides an easy explanation for the difference in our results, but rather that the difference is more likely due to the combination of many subtle choices.

\section{Conclusion}
The hunt for exomoons remains exciting and volatile. Our analysis of the exomoon candidate Kepler-1625b I shows that the evidence in favor of the moon is highly sensitive to details of the data reduction and modeling. Given that we achieve an excellent fit to the light curve with a no-moon model, and that the addition of a moon does not significantly improve the fit quality, we suggest that future allocation of follow-up resources for exomoons should prioritize alternate targets.

>>>>

\begin{abstract}
    Observations of the Kepler-1625 system with the Kepler and Hubble Space Telescopes have suggested the presence of a candidate exomoon, Kepler-1625b I, a Neptune-radius satellite orbiting a long-period Jovian planet. Here we present a new analysis of the Hubble observations, using an independent data reduction pipeline. We find that the transit light curve is well fit with a planet-only model, with a best-fit $\chi^2_\nu$ equal to $1.01$. The addition of a moon does not significantly improve the fit quality. We compare our results directly with the original light curve from \cite{teachey18b}, and find that we obtain a better fit to the data using a model with fewer free parameters (no moon).  We discuss possible sources for the discrepancy in our results, and conclude that the lunar transit signal found by \cite{teachey18b} was likely an artifact of the data reduction.  This finding highlights the need to develop independent pipelines to confirm results that push the limits of measurement precision.
\end{abstract}

@@

We found that our results were unchanged when we used the exponential systematics model from TK18 (described in \S\,\ref{sec:sys}). For this case, the moon model is also strongly disfavored ($\Delta\rm{BIC} = 32.2$). The fit rms is within 5 ppm of the XY decorrelation model.

\subsection{Comparison with Teachey \& Kipping (2018)}
TK18 found evidence for the transit of a Neptune-size moon in their analysis of the HST data, in contrast to the findings presented here.  To compare our results with theirs, we fit the TK18 light curve directly.  We fit the astrophysical signal with both the no-moon and models, and used the exponential systematics model.  Figure\,\ref{fig:bestfit} shows the best fit models.

%For the systematics, we tested the XY decorrelation model described above, but found that it did not perform as well as the systematics models presented in TK18 (10\% higher rms and time-correlated residuals).  We therefore opted to use a TK18 systematics model to enable a fair comparison between our results. We obtained the best fit using an exponential trend in time, a constant offset after orbit 14 (where the guide stars are reacquired), and the non-parameteric orbit-long ramp model.  This model has one more free parameter than the XY decorrelation model given in Equation~\ref{eq:sys}.

Similar to the findings of TK18, the moon model improved the fit quality by a $\Delta\chi^2 = 29.9$. Notably, however, the moon model fit to the TK18 data does not perform better than the no-moon model fit to our new data (rms of 362 versus 356 ppm), even with the additional seven free parameters.

@@

\section{Discussion \& Conclusions}
A natural question arising from our analysis is the source of the discrepancy between TK18 and the new results presented here. We find that with our new data, there is strong evidence against the moon ($\Delta\rm{BIC} > 25$), even when we use a comparable model to TK18 (a 6-parameter moon model and an exponential fit to the visit-long trend).

If the model is not the source of the difference, it must be the extracted transit light curves themselves.  Figure\,\ref{fig:raw} shows a direct comparison of the light curves. The count rate we measure is $2.46 -2.74\%$ lower than the TK18 light curve, and there is a small bump in the difference between the two data sets near the location of the moon transit identified in the TK18 data (see the bottom panel). This bump may be the source of the moon feature reported in TK18.

 We explored several modifications to our pipeline to attempt to reproduce the TK18 data reduction. These included rotating the image by 0.5 degrees, using the same aperture as TK18 to extract the spectrum, and scaling the master sky flat for the background subtraction (rather than just subtracting the median). None of these modifications had a significant effect on our results.

@@

%In addition to these differences in data reduction, we also used an alternative detrending model. We opted for a decorrelation against X and Y position, motivated by known variation in the sensitivity of individual pixels on the detector and the presence of small shifts in the position of the spectrum (see Figure\,\ref{fig:shifts}). We applied this detrending model to the TK18 data, and obtained a poor fit to the light curve compared to exponential detrending (for the moon model with XY decorrelation, the rms was 404 ppm). We did not recover a large moon in this fit (the best fit was $r_\mathrm{moon} = 1.2R_\oplus$), but there was correlated noise present in the data that precluded a robust measurement of the uncertainty on the moon radius for this model. The poor quality of the fit points toward the extracted light curve itself as the culprit for the difference in our results, but as we discuss, we could not identify a specific step in the data reduction that is responsible.

%\section{Conclusion}
%The hunt for exomoons remains exciting and volatile. Our analysis of the exomoon candidate Kepler-1625b I shows that the evidence in favor of the moon is highly sensitive to details of the data reduction. Given that we achieve an excellent fit to the light curve with a no-moon model, and that the addition of a moon does not significantly improve the fit quality, we suggest that future allocation of follow-up resources for exomoons should prioritize alternate targets.
